
\coursecode{MA 103}
\coursename{Mathematics-I}
\credit{4}
\lecture{3}
\tutorial{1}
\lab{0}
\section{\courseinfo}

\unit{I}
Sets and their representations, operations on sets, Cartesian product of sets, relations, functions, algebraic and transcendental function, Principle of Mathematical induction, Solution of quadratic equations. Permutation and Combination, Binomial Theorem.

\unit{II}
Straight Lines, slope of a line and angle between two lines, various forms of equations of a line: parallel to axes, point-slope form, slope-intercept form, two point forms, intercepts form and normal form, general equation of a line, distance of a point from a line. Introductory idea of Ellipse, Circle, parabola and Hyperbola.
 
\unit{III}
 Trigonometric functions, positive and negative angles, Measuring angle in radians \& in degree and conversion from one measure to another. Definition of trigonometric  with the help of unit circle. Trigonometric identities. Complex numbers, algebraic properties of complex numbers, Argand plane and polar representation of complex numbers, Fundamental theorem of algebra.
 
\unit{IV}
Vectors and scalars, magnitude and direction of a vector, direction cosines (and ratios) of vectors, Types of vectors, position vector of a point, negative of a vector, components of a vector, addition of vectors, multiplication of a vector by a scalar, position vector of a point dividing a line segment in a given ratio. Scalar product of vectors, projection of a vector on a line, vector product of vectors.
 
\unit{V}
 Matrices and determinants: definition of a matrix, various types of matrices, addition, subtraction, multiplication of matrices, inverse of matrix, determinant of matrices, expansion of determinant, properties of determinants, solution of linear system of equations, Cramer rule.

\paragraph{Textbook}
\emph{Mathematics Part I and Part II - Textbook for Class XI and XII}, NCERT.

\begin{thebibliography}{}
\let\clearpage\relax
\bibitem{1} H. S. Hall and S.R. Knight, \emph{Higher Algebra}. Arihant, 2010.
\bibitem{2} J. Stewart, \emph{Calculus}, Cengage Learning, Sixth Edition.

\end{thebibliography}














