\coursecode{MA 108}
\coursename{Probability and Statistical Methods}
\credit{2}
\lecture{2}
\tutorial{0}
\lab{0}
\section{\courseinfo}
%M. Tech SOBT-3 years (II semester)
\unit{I}
Statistical Methods: Classification of data, Concepts of statistical population and sample from a population, representation of data in form of tables and diagrams, Histogram and frequency polygon, cumulative frequency distributions Ogive and Pie charts, Bivariate data-scatter diagram, principle of least squares: fitting of polynomials and exponential curves, Correlation and regression, Covariance, Measures of location (or central tendency) and dispersion, moments, measures of skewness and kurtosis and Introduction to time series analysis.

\unit{II}
Probability Theory: Random experiments, sample point and sample space, event, algebra of events, Definition of Probability – classical and relative frequency approach to probability: theorems on probability, conditional probability, independent events, Bayes theorem and its applications. 

\unit{III}
Random variables (discrete and continuous): Distribution function, Probability mass and density functions, Expectation, Moments, Moment generating function, Chebyshev’s inequality. Special distributions: Bernoulli, Binomial, Uniform, Normal, distributions, Independence of random variables.

\unit{IV}
Functions of survival time, survival distributions and their applications, Reliability and hazard rate function, Stochastic epidemic models: Simple and general epidemic models, Markov chain and applications of Markov Chains in Bio-informatics.


\begin{thebibliography}{99}
        \bibitem{k} Goon A.M., Gupta M.K. and Dasgupta B. (2005): \emph{Fundamentals of Statistics}, Vol. I, 8th  Edn. World Press, Kolkata. 
        \bibitem{k} Gupta, S.C. and Kapoor, V.K. (2007): \emph{Fundamentals of Mathematical Statistics}, 11th  Edn.,(Reprint), Sultan Chand and Sons. 
      \bibitem{k} Rohatgi, V. K. and Saleh, A. K. Md. E. (2009): \emph{An Introduction to Probability and Statistics}, 2Nd Edn. (Reprint). John Wiley and Sons.
        \bibitem{k} Chiang, C.L. (1968): \emph{Introduction to Stochastic Processes in Bio Statistics}. John Wiley and Sons.

\end{thebibliography}


