\coursecode{MA 205}
\coursename{Quantitative Techniques in Food Technology}
\credit{3}
\lecture{3}
\tutorial{0}
\lab{0}
\section{\courseinfo}

\unit{I}
Collection of Data: Classification and Tabulation, Variable and Attribute, Primary data and secondary data, Population and sample, Types of charts and diagrams, averages or measures of central tendency, Measures of central tendancy, relation between mean, median and mode.

\unit{II}
Frequency distribution: observation, frequency simple series and frequency distribution, cumulative frequency distribution, relative frequency distributions, diagrammatic representation of frequency distribution, frequency curve.

\unit{III}
Measures of dispersion: Range, meandeviation, standard deviation, relative measures of dispersion, lorenz curve, moments, skewness and kurtosis, beta coefficients and gamma-coefficients, curve fitting and method of least squares.

\unit{IV}
Concepts of correlation and regression, bivariate data, bivariate frequency distribution, scatter diagram, covariance, correlation coefficient, properties of correlation coefficients, regression, properties of linear regressions, expalined variation and unexplained variation, Random Variable, Discrete Random Variable, Probability Function, Continuous Random Variable, Probability Density Function, Discrete Distributions, Continuous Distributions.



\paragraph{Textbook}
J. N. Kapoor, Mathematical Statistics, S. Chand Publication, 2004.
\begin{thebibliography}{}
\let\clearpage\relax
\bibitem{1}     V. K Rohatgi, \emph{An Introduction to Probability Theory and Mathematical Statistics}, John Wiely \& Sons 1976.
\bibitem{4}     N. G. Das, \emph{Statistical Methods}, McGraw Hill Education (India), Private Limited 2013.
\bibitem{5}     Marylees Miller, John E. Freund, Irwin Miller, \emph{Mathematical Statistics: With Applications,} Prentice Hall 2003.
\end{thebibliography}













