\coursecode{MA 202}
\coursename{Numerical and Statistical Analysis}
\credit{4}
\lecture{3}
\tutorial{1}
\lab{0}
\section{\courseinfo}


\unit{I} Zeroes of transcendental and polynomials, Bisection method, Regular-falsi method and Newton-Raphson method, Secant Method, Rate of convergence of above methods, Error Analysis.

\unit{II} 
Interpolation, Finite differences, difference tables, Newton'’s forward and backward interpolation, Divided differences, Lagrange'’s and Newton'’s divided difference formula for unequal intervals. Numerical differentiation, Numerical integration, Newton-Cote quadrature formula, Trapezoidal, Simpson'’s one third and three-eight rules. 

\unit{III}
Solution of system of linear equations, Gauss-Seidal method, Gauss-Jordan Method, Crout method, Jacobi Iteration method.
Numerical Solutions of Ordinary Differential Equations and Partial Differential Equations: Picard's Method, Euler's Method, Modified Euler's Method, Runge-Kutta
method. Numerical solutions of parabolic and elliptic partial differential equations, Introduction to finite element method.

\unit{IV}
Probability, Conditional Probability, Bayes' Theorem, Random variable, Probability distributions,Mean and Variance of the distribution, Probability density function, Binomial distributions, Poisson distributions and Normal distributions.

\unit{V}
Mathematical Expectations, Moments, Moment generating functions, Skewness, Kurtosis, Curve fitting, Method of least squares, Fitting of straight lines, Polynomials, Exponential curves, Correlation, Regression analysis: Linear, Nonlinear and multiple. Random sampling,Sample mean, Sample variance, Estimation of parameters, Maximum likelihood method, Confidence intervals, Testing of Hypotheses:in the case of normal distributions, Goodness of Fit, Chi-square test, t-test, Analysis of variance.
 
\paragraph{Textbook}
 E. Kreyszig, \emph{Advanced Engineering Mathematics}, John Wiley and Sons, 2005.


\begin{thebibliography}{}
\let\clearpage\relax
%\bibitem{1} E. Kreyszig, \emph{Advanced Engineering Mathematics}, John Wiley and Sons, 2005.
\bibitem{2} R. S. Gupta, \emph{Elements of Numerical Analysis}, Macmillan, 2009
\bibitem{3} C.F. Gerald and P.O. Wheatley, \emph {Applied Numerical Analysis} 7th Edit., Addison-Wesley, Boston, 2003.
\bibitem{4} R. A. Johnson, \emph{Miller \& Freund's Probability and Statistics for Engineerings} Prearson, 2001
\bibitem{4} S. M. Ross, \emph{Introduction to Probability and Statistics for Engineers \& Scientists} Elsevier, Academic press, 2004.

\end{thebibliography}














