\coursecode{MA 203}
\coursename{Quantative Techniques in Biotechnology}
\credit{4}
\lecture{3}
\tutorial{1}
\lab{0}
\section{\courseinfo}

\unit{I}
Frequency Distribution, Graphic representation of a frequency distribution, Measures of Central tendency, Moments, moment generating functions, Central Moments, Non- Central Moments, Measures of Dispersion, Measures of Skewness, Measures of Kurtosis 

\unit{II}
Curve fitting, Method of Least squares, Fitting of Straight lines, Polynomials, Exponential Curves, Linear Correlation, Correlation Coefficient, Properties of Correlation Coefficient, Rank Correlation Coefficient, Regression Analysis: Linear \& Non linear. Time series and forecasting.

\unit{III}
Introduction, Experiments, Outcomes, Events, Definition of Probability, Conditional Probability, Independent Events, Total Probability, Bayes' Theorem.

\unit{IV}
Random Variable, Discrete Random Variable, Probability Function, Continuous Random Variable, Probability Density Function, Discrete Distributions, Continuous Distributions, Mathematical Expectation, properties of Expectation, Two Dimensional Random Variables, Joint Probability Density Functions, Marginal Probability Distribution, Conditional Probability Distribution.

\unit{V}
Binomial Distribution, Poisson Distribution, Normal Distribution
Sampling theory, Tests of significations, Chi-square test, t-test, Analysis of variance (one way), 
Statistical quality control methods, Control charts, R, p, np, and c charts.

\paragraph{Textbook}
J. N. Kapoor, Mathematical Statistics, S. Chand Publication, 2004.
\begin{thebibliography}{}
\let\clearpage\relax
\bibitem{1}     V. K Rohatgi, \emph{An Introduction to Probability Theory and Mathematical Statistics}, John Wiely \& Sons 1976.
\bibitem{4}     Marylees Miller, John E. Freund, Irwin Miller, \emph{Mathematical Statistics: With Applications,} Prentice Hall 2003.
\bibitem{5} Gupta, S.C. and Kapoor, V.K. (2007): \emph{Fundamentals of Mathematical Statistics}, 11th  Edn.,(Reprint), Sultan Chand and Sons.
\bibitem{5} S. M. Ross, \emph{Introduction to Probability and Statistics for Engineers \& Scientists} Elsevier, Academic press, 2004.

\end{thebibliography}













