\coursecode{MA 204}
\coursename{Numerical Methods}
\credit{3}
\lecture{2}
\tutorial{0}
\lab{2}
\section{\courseinfo}

Note: This course may be offered for 3 Credits L-T-P (2-0-2). In this case Unit-IV and V will be removed and Implementation of methods discussed in remaining will also be done using C++/Python/MATLAB.


\unit{I} Zeroes of transcendental and polynomials, Bisection method, Regular-falsi method and Newton-Raphson method, Secant Method, Rate of convergence of above methods, Error Analysis.

\unit{II} 
Interpolation, Finite differences, difference tables, Newton'’s forward and backward interpolation, Divided differences, Lagrange'’s and Newton'’s divided difference formula for unequal intervals. Numerical differentiation, Numerical integration, Newton-Cote quadrature formula, Trapezoidal, Simpson'’s one third and three-eight rules. 

\unit{III}
Solution of system of linear equations, Gauss-Seidal method, Gauss-Jordan Method, Crout method, Jacobi Iteration method.
Numerical Solutions of Ordinary Differential Equations and Partial Differential Equations: Picard's Method, Euler's Method, Modified Euler's Method, Runge-Kutta
method. Numerical solutions of parabolic and elliptic partial differential equations, Introduction to finite element method.


\paragraph{Textbook}
E. Kreyszig, \emph{Advanced Engineering Mathematics}, John Wiley and Sons, 2005.


\begin{thebibliography}{}
	\let\clearpage\relax
	%\bibitem{1} E. Kreyszig, \emph{Advanced Engineering Mathematics}, John Wiley and Sons, 2005.
	\bibitem{2} R. S. Gupta, \emph{Elements of Numerical Analysis}, Macmillan, 2009
	\bibitem{3} C.F. Gerald and P.O. Wheatley, \emph {Applied Numerical Analysis} 7th Edit., Addison-Wesley, Boston, 2003.
	
\end{thebibliography}
