\coursecode{MA 201}
\coursename{Engineering Mathematics-III}
\credit{4}
\lecture{3}
\tutorial{1}
\lab{0}
\section{\courseinfo}
\unit{I}
Introduction: linear, nonlinear, order, degree, and types of solutions of ODE, exact differential equations, integrating factors, linear equation, Bernoulli equations, Picard's existence and uniqueness theorem for $dy/dx=f(x,y)$ (without proof). ODE of first order not of first degree, Solvable for $x$, Solvable for $y$, Solvable for $p$, Clairaut's form.

\unit{II}
Linear differential equations of nth order with constant coefficients, homogeneous and non-homogeneous linear differential equations, simultaneous linear differential equations, Euler -Cauchy linear differential equations, method of variation of parameters, %applications to engineering problems (Motion of a particle in resisting medium, simple harmonic motion, electric circuit  problem).
Power series method for ODE (Frobenious Method).
(\emph{Optional Topics: Legendre function and Bessel function and their properties.})

\unit{III}
Laplace Transform: Existence theorem, Properties of Laplace transform, Inverse Laplace transform, Unit step function, Dirac delta function, Laplace transform of periodic functions, Convolution theorem, Application to solve linear and simultaneous differential equations. 

\unit{IV}
Periodic functions, Trigonometric series, Fourier series of period $2\pi$, Euler's formulae, Functions having arbitrary period, Change of interval, Even and odd functions, Half range sine and cosine Fourier series. 

\unit{V} PDEs and its Applications: 
Linear partial differential equations with constant coefficients (Operator method only). Classifications of $2^{nd}$ order PDE. Method of separation of variables for solving partial differential equations, Its applications to solve  Heat conduction equation, Wave equation, steady state heat equation (Laplace equation) through Fourier series.

\paragraph{Textbook}
R. K. Jain and S.R.K Iyengar, \emph{Advanced Engineering Mathematics}, Narosa Publications.

\begin{thebibliography}{99}
\let\clearpage\relax
\bibitem{2} G. F. Simmons, \emph{Differential Equations}. Tata Mc Graw-Hill Publishing Company Ltd., 1981.
\bibitem{3} E. Kreyszig, \emph{Advanced Engineering Mathematics}. John Wiley and Sons, 2005.
\bibitem{4} W. E. Boyce and R.C. DiPrima, \emph{Elementary differential equations and Boundary Value Problems}. John Wiley and Sons, 8th Edition, 2010.

\end{thebibliography}














