\coursecode{MA 402}
\coursename{Modeling and Simulation}
\credit{4}
\lecture{3}
\tutorial{1}
\lab{0}
\section{\courseinfo}

%[M. Tech ICT (II semester- specialization: VLSI, ISR, SE, WCN)]

\unit{I}
Introduction to Simulation and Modeling, Simulation process, Advantages and disadvantages of simulation techniques, Limitations of simulation techniques, Comparison of simulation and analytical methods, fixed time-step vs even to even model, Analog vs digital Simulation. 

\unit{II}
Simulation of continuous systems, Simulation of water reservoir
system, Simulation of a servo system, simulation of an autopilot.


\unit{III}
Discrete system simulation, Queuing Models, Characteristics of
Queuing Models, Behaviour of arrivals, Pattern of arrival at the
system, The arrival time distribution, Queuing process, Queuing
discipline, Service process, Distribution of service time, Performance
measures of a queuing system, Classification of queuing
models, Single server queuing models and its simulation. Inventory model, Deterministic inventory models with and without
shortage cost model, Simulation of Inventory models.
 

\unit{IV}
Monte- Carlo simulation, Real time simulation, Hybrid simulation,
Distributed Log models, Cobweb model, Generation of random
numbers, Test for randomness, confidence interval, Statistical
methods. CPM and PERT networks, Characteristics, Critical path computation.


\unit{V}
System dynamics, System dynamic diagrams, Exponential growth models, Exponential decay models, Modified exponential growth models, Verification and Validation of Simulation Models, Techniques for verification and validation of model, Time Series Approach.


\begin{thebibliography}{99}
        \bibitem{1}     Law M, Kelton W, \emph{Simulation Modeling and Analysis}. McGraw-Hill, New York, NY, 2000.
        \bibitem{2}     Khoshnevis B, \emph{Systems Simulation - Implementations in EZSIM}. McGraw-Hill, New York, NY, 1992.
        \bibitem{3}     Geoftrey Gordon, \emph{ System Simulation}, PHI.
        \bibitem{4}     Jerry Banks, John S. C Barry L. Nelson David M. Nicol, \emph{Discrete Event System
        Simulation}, Pearson Education.
       % \bibitem{1}     Averill M. Law, W. David Kelton, \emph{System Modeling and simulation and Analysis},
        TMH.
\end{thebibliography}
