\coursecode{MA 415}
\coursename{Biostatistics}
\credit{3}
\lecture{3}
\tutorial{0}
\lab{0}
\section{\courseinfo}

\textbf{Unit I}

Descriptive Statistics: Application of statistics in managerial decision-making, Collection of data, Tabulation and graphic presentation of data, Measures of Central tendency (Mean, Median and Mode), Measuring the variation in Data, Standard Deviation, Population Variance, Sample Variance, Significance of Standard Deviation; Percentiles, Quartiles, Skewness. Distribution of sample means, standard error and confidence interval.

\textbf{Unit II }

Correlation and Regression: Correlation: Meaning and uses, Methods of correlation, Regression: Meaning and uses, Regression equations, Time series: Concepts, Components of time series, Measurement of trend. Scatter plot, multiple regressions.

\textbf{Unit III}

Probability: Permutation and combination, Sample space and events, probability concepts, addition theorem, multiplication theorem, Conditional probability.

% Central limit theorem; Inference; Hypothesis; Critical region and error probabilities; Tests for proportion; Equality of proportions; equality of means of normal populations (variance known, variance unknown); P-value of the statistic; Confidence limits.

\textbf{Unit IV}

Probability Distributions: Discrete Random Variables, Continuous Random Variables, Expected Value, Variance, Introduction to one way and two-way analysis of variance; Data transformations. Binomial Distribution, Poisson distribution, Normal Distribution.

%Chi-squared Distribution, Students-t Distribution, F-test, Z-test. Hypothesis testing: null and alternative hypotheses, decision criteria, critical values, type I and type II errors, one tailed and two tailed tests.  

\begin{thebibliography}{99}
        \bibitem{1}     Wayne W. Daniel, Biostatistics : \emph{Basic Concepts and Methodology for the Health Sciences}, 9th edition, Wiley India Pvt Ltd, 2008        \bibitem{3}     Sheldon Ross, \emph{A First Course in Probabilit}y, 9th Edition,
, 2014.
        \bibitem{2}     Edward Batschelet,  \emph{Introduction to Mathematics for Life Scientist}s, 3rd Edition, Springer-Verlag, 1992.
\bibitem{aa} S.P. Gordon and F.S. Gordon, \emph{Contemporary Statistics:A Computer Approach}, McGraw-Hill Publishing Company, New York, 1994. 
           
\end{thebibliography}
