\coursecode{MA 404}
\coursename{Topology}
\lecture{4}
\tutorial{1}
\lab{0}
\section{\courseinfo}

Prerequisite: Basic sets theory, Cartesian Products, Relations, Finite Sets, Countable and uncountable sets, Infinite Sets and the Axiom of Choice, Well ordering, Zorn’s Lemma and their equivalence.. 

Metric Spaces, Neighbourhoods and Continuity, Topological spaces, Basis for a topology, Order topology, Product topology, Subspace topology, Closed sets, Limit points, Continuous functions, Homeomorphism, Quotient topology, Metric topology, Connected spaces, Components, Compact spaces, Local compactness, Path-connected spaces, Path-connected components, Countability axioms, Separation axioms, Hausdorff spaces, Normal spaces, Regular spaces, Urysohn lemma, Urysohn metrization theorem, Tietze extension theorem, Tychonoff theorem. One-point compactification, Stone Cech compactification.
Supplementary Topics: Topological Groups, Nets

Text Book

J. R. Munkres, Topology, 2nd ed., Pearson, 2006

References Books 

[1] S. Willard, General Topology, Addison – Wesley. 

[2] C. Adams and R. Franzosa, Introduction to Topology: Pure and Applied, Pearson, 2000.

\coursecode{MA 406}
\coursename{Operations Research}
\section{\courseinfo}

Prerequisite: Linear Algebra.  

Basics of Linear Programming Problem, Convex sets and their properties, Duality in linear programming, Complementary slackness theorem, Game Theory, Linear programming and matrix game equivalence, Sequencing Problem, Processing n Jobs through two, three, and machines, Replacement Problems, Deterioration with time, cost increasing with time, money value into consideration, Critical Path Determination by CPM and PERT, Goal Programming Inventory control models, Dynamic Programming.

Supplementary Topics: Sensitivity Analysis of LPP, Transportation problem.

Text Book

J.C. Pant, Introduction to optimization: Operations Research, Jain Brothers, New Delhi, 2002.

References Books

[1] A. Ravindran, D. Phillips, and J. Solberg, Operations Research: Principles and Practice, 2nd Ed., Wiley India,2007.

[2] H. A. Taha, Operations Research: An Introduction, Pearson Prentice Hall, 2004.



\coursecode{MA 408}
\coursename{Partial Differential Equations}
\section{\courseinfo}

Prerequisite: Analysis.  

Introduction, Classification to first order PDEs, Mathematical models leading to PDEs, Method of Characteristics, Scalar conservation laws and linear transport equation, Traffic dynamics, The Rankine-Hugoniot condition, The entropy condition, The Riemann problem, The Cauchy problem, Classification of second order equations: canonical forms of Hyperbolic, Parabolic, elliptic equations. One dimensional wave equation and De’Alembert’s method, Domain of dependence and region of influence, The Cauchy problem for the non homogeneous wave equation, Methods of separation of variables for heat, Laplace and wave equations, The energy method and uniqueness, Sturm–Liouville problems and eigen function expansions, Elliptic equations: maximum principle, applications of the maximum principle, Green’s identities, The maximum principle for the heat equation, Poisson’s formula, Laplace equation: mean value property, weak and strong maximum principles, Green`s function, Poisson`s formula, Dirichlet`s principle, Harnack’s inequality and Liouville’s theorem. Green’s functions and integral representations

Text Book

Y. Pinchover and J. Rubinstien, Introduction to Partial Differential Equations, Cambridge University Press.

References Books 

[1] Sadro Salsa, Partial Differential Equations in action, Springer-verlag Italia, Milano 2008.

  	[2] L.C. Evans, Partial Differential Equations, Graduate Studies in Mathematics, Vol. 19, American Mathematical Society, 1998.

[3] Walter A Strauss, Partial Differential Equations, 

[4] I.N. Sneddon, Elements of Partial Differential Equations, Mc-Graw Hill .

\coursecode{MA 410}
\coursename{Complex Analysis}
\section{\courseinfo}
Prerequisite: Basic real analysis, complex number system, Complex plane.  

Limits, Continuity, and differentiability of complex variable functions, Comparison between differentiability of real valued function and complex valued functions. Analytic functions, Necessary and Sufficient condition, The Cauchy-Riemann equations, Entire functions, Harmonic function and Harmonic conjugate function, Power series and radius of convergence of power series. Conformal Mapping, Cross ratio, Bilinear Transformation, Integration of complex-valued functions and differential forms along a piecewise differentiable path, Cauchy theorem, Cauchy integral formula, Morera’s Theorem, Residues, Cauchy’s Residue theorem, Evaluation of integrals by residue method, Taylor’s expansion of holomorphic functions, Liouville’s theorem, fundamental theorem of algebra, Maximum Modulus theorem; Schwarz’ lemma, Laurents’s expansion of a holomorphic function in an annulus, singularities of a function, argument principle, Rouche’s Theorem.

Supplementary Topics: Stereographic projection, branch points, elementary functions: polynomial, rational, exponential and hyperbolic functions.

Text Book

H. Silverman, Complex Variables.

References Books 

[1] L. V.  Ahlfors, Complex Analysis (Third Edition), McGraw-Hill (International Editions).

[2] J. B. Conway : Functions of one complex variable

[3] Brown and Churchill, Complex Analysis, McGraw Hill.

[4] Moore, Haddock, Complex Analysis

\coursecode{MA 412}
\coursename{Measure and Integration}
\section{\courseinfo}

Prerequisite: Review of Riemann Integral, Riemann-Stieltjes Integral.

Outer Measures, Carathodory Extension, Lebesgue Measure; Lebesgue Outer Measure; Lebesgue Measurable Sets. Measure on an arbitrary sigma-Algebra; Measurable Functions; Integral of a Simple Measurable Function; Integral of Positive Measurable Functions. Lebesgue’s Monotone Convergence Theorem; Integrability; Dominated Convergence Theorem; Lp-Spaces. Absolute Continuous Function on R, Differentiation and Fundamental theorem for Lebesgue integration. Measure on Product Spaces;  Fubini’s theorem. 

Supplementary Topics: Complex measures, Radon-Nikodyn theorem.

Text Book: G. D. Barra, Measure and Integration, Wiley Eastern, 1981

References Books 

[1] G. D. Barra, Measure and Integration, Wiley Eastern, 1981. 

[2] Terence Tao, An Introduction to Measure Theory 

[3] W. Rudin, Real and Complex Analysis, Third edition, McGraw-Hill, International Editions, 1987.

[4] I. K. Rana, An Introduction to Measure and Integration, Second Edition, Narosa, 2005. 

[5] D. L. Cohn, Measure Theory , Birkhauser, 1997. 

[6] H. L. Royden, Real Analysis, Third edition, Prentice-Hall of India, 1995.
