\coursecode{MA 501}
\coursename{Functional Analysis}
\section{\courseinfo}

Prerequisite: Metric Spaces, Holder’s and Minkowski’s inequalities, Completeness, Compactness, Heine-Borel and Bolzano-Weierstrass Theorems in $R^n$, Continuity, Uniform continuity, Urysohn lemma, Tietze extension theorem.

Normed spaces, Banach spaces, Riesz Lemma, Bounded linear operator, Ascoli theorem, Hahn-Banach extension and separation theorems, Dual of classical spaces $l^p$ , $L^p$,  Banach contraction mapping theorem, Uniform boundedness principle, Open mapping and closed graph theorems, weak convergence, Inner Product space, Hilbert space, Orthonormal sets, Riesz representation theorem, Fourier series -$L^2$ theory, Bounded linear operators on Hilbert spaces. Spectral theorem for compact self-adjoint operators.

Supplementary Topics: Weak and weak-∗topologies, Banach Alaoglu theorem, Spectral theorem for general self-adjoint and normal operators. Reisz representation theorem, dual of $CO(X)$, $X$ a locally compact space, Gelfand theory. Unbounded operators: definition and examples.

Text Book

G. F. Simmons, Introduction to Topology Modern Analysis

References Books

[1] B. V. Limaye: Functional Analysis, Revised 2nd edition,New Age International Limited, 1996.

[2] I. J. Maddox; Elements of Functional Analysis, Cambridge University Press.

[3] E. Kreyszig; Functional  Analysis, John Wiley.

\coursecode{MA 503}
\coursename{Integral Equations and Calculus of Variation}
\section{\courseinfo}

Prerequisite: Basic differential equation. 

Introduction and formulation of integral equation, Classification of Integral Equations., Conversion of IVP(ODE) to Volterra equation, Conversion of BVP(ODE) to Fredholm equation, Existence and uniqueness of solutions using fixed-point theorems in case of Linear and nonlinear Volterra and Fredholm integral equations using Banach Fixed point theorem, Solutions with separable kernels. Characteristic numbers and eigenfunctions, resolvent kernel. Greens function. Calculus of Variations: Variation of a functional, Euler-Lagrange equation, Necessary and sufficient conditions for extreme.

Text Book

L. Debnath, D. Bhatta, Integral transforms and their applications, 2nd ed, Chapman \& Hall/CRC, 2007

References Books 

[1] O. Bolza, Lectures on the Calculus of Variations, Dover Publication, New York, 1961.

[2] J. A. Cochran, The Analysis of Linear Integral Equations, McGraw-Hill, 1972

[3] Francis B. Hildebrand, Methods of Applied Mathematics, Dover, New York, 2012

[4] G. F. Roach, Greens Functions, Cambridge University Press, 1995.
